Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy

### 2020-09-30 manuscript/manuscript.Rmd
Add survey text
Suparna edits

---
title: "Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy"
short-title: "Who Cares About Crackdowns?"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
date: "September 24, 2020"
published: Working paper.
git-repo: https://example.com
abstract: >-
  Over the past two decades, international non-governmental organizations (INGOs) have faced increasing legal scrutiny and repression abroad. As a result, official aid to INGOs in repressive countries has reduced. NGOs thus need access to a diverse array of funds and may need to increase fundraising from private donors. Which demographic and experiential factors influence the decision of individuals to donate to legally besieged NGOs? We investigate this question using a market simulation based on a nationally representative conjoint experiment conducted on individuals with a history of charitable giving. We find that trust in political institutions and charities is enormously influential in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to human rights INGOs with friendly relationships with host governments, and that support hold steady if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do not face criticism or crackdown abroad. Our findings have practical implications for INGOs that aim to raise funds from the general public amid closing civic space. 
reference-section-title: References
toc: false
mainfont: Lora
sansfont: IBM Plex Sans
---

@@

As a result of these legal crackdowns, foreign aid channeled through NGOs has decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While this has negatively impacted funding flows from public organizations like United States Agency for International Development (USAID) and Organisation for Economic Cooperation and Development (OECD) philanthropy from private donors and foundations appears to not be as adversely affected [@McGill:2018]. Given this trend of closing civic space, INGOs need to maintain access to multiple sources of funding, including individual donors.

Individual private donations are a critical source of INGO funding. Total private philanthropy in the United States has grown substantially over time, reaching more than \$400 billion in 2018 (@fig:giving-aggregate). In 2018, private giving to NGOs working specifically in international affairs reached \$22.08 billion, and international affairs-focused organizations are one of only two categories of organizations that have continued experiencing substantial growth in giving [@givingusa:2019].[^intlaffairs] Most of this growth is driven by individuals making small-scale or micro-donations.[^smalldonors] In 2017, 70% of donations to international affairs NGOs came from individuals, rising to 86% in 2018.  But relying on individual donors means that NGOs must consider individual preferences when fundraising. Which individual donor characteristics and attributes are most salient for INGOs interested in maximizing philanthropy at a time when they are facing increasingly hostile environments abroad? What kinds of donors are more or less likely to consider an INGO's host country political environment when making the decision to donate?

@@

Existing scholarship on the determinants of individual-level philanthropy to domestic causes is robust across a variety of contexts. But the dynamics of *international* giving to NGOs are under-examined in the literature on individual donor attributes and preferences.[^exceptions] Motivating individuals to donate to international causes is hard as the number of recipients are large and far removed from the donor [@CasaleBaumann:2015, 100]. And research on the determinants of philanthropy for international causes—particularly towards organizations working in challenging environments—is still in its infancy due to the absence of expansive and high-quality data [@greenhill2013]. While previous work has looked at how organizational attributes of nonprofits affect donors' willingness to donate internationally [@ChaudhryHeiss:2020], we have yet to uncover demographic traits and experiences that motivate or deter individuals to donate during shrinking civic space.

We explore this question using a simulated market for philanthropic donations, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing to generate dozens of distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that specific individual characteristics help shape donor preferences for donating to international NGOs, and that an INGO's target issue area and its relationship with host government interact strongly with these characteristics. Personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with more contentious relationships with host governments, preferring instead to donate to NGOs with friendly host-country relationships and working on less contentious issues like emergency response and refugee relief. Personas with higher trust are more comfortable donating to legall besieged INGOs.

This study contributes to both to scholarship and practice in public policy, nonprofits, and philanthropy. Individual giving abroad can help bridge socioeconomic disparities worldwide, especially given the stark inequality in global distribution of income. Knowledge of demographic traits and its impact on funding can help nonprofits craft more effective and targeted marketing strategies [@Rajan:2009, 414]. Most NGOs working internationally have previously relied on government and foundation funding. But with shrinking civic space, these groups may need to readjust their fundraising strategies towards individuals. Framing is important in this regard—how people process information affects their willingness to donate.[^framing] However, existing research is unclear about which frames are effective when making appeals to individual donors considering international philanthropy. Since our study looks at multiple organizational and host country factors, our results can help NGOs frame their appeals.

@@

Below, we first examine the state of the existing literature on the key drivers of individual giving. We then describe our survey experiment and simulation methods and explore the results. We conclude with a discussion of the implications our findings have on general INGO funding in the age of constrained civil society space.

@@

Foreign aid to civil society—and NGOs in particular—has steadily increased over time. Many government and foundation donors perceive NGOs as nimbler and less bureaucratic than governments. In countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency, donors specifically seek out NGOs for projects [@Dietrich2013].

@@

The advent of funding restrictions on NGOs globally and a subsequent decline in government and foundation funding has made the role and value of individual philanthropy in sustaining NGOs even clearer.[^phildef] NGOs need access to a diverse array of funds and individual donations to international affairs organizations have steadily increased over the past two decades. Moreover, in the United States donations from individuals have far outstripped foundation and corporate giving. In 2018, individual donations amounted to \$292 billion, compared to \$76 billion from foundations and \$20 billion from corporations [@NPT]. Given that a majority of research is restricted to elite high net-worth donors[^highnetworth] [@USTrust2014], these figures show that it is imperative to understand motivations and attributes of small-scale donors as well.

@@

A vast literature on philanthropy looks at organizational-level factors that motivate individuals to donate [@TBPrakash:2017]. However, donation decisions are not determined solely by a recipient organization's characteristics. Donors have personality traits, preferences, and experiences that make them more or less likely to engage in philanthropy. Substantial research has looked at the determinants of donating to local—rather than international—causes, including four main attributes: social capital, the role of higher education, levels of religiosity, wealth, and national group attachment.

@@

Higher education also influences charitable-giving decisions as it fosters prosocial motivations. Education brings people into social networks that entail a higher level of solicitations, which is vital in giving to international causes [@BekkersWiepking:2011]. Education also helps determine the kinds of causes that people donate to globally. @Micklewright:2009 show that higher education matters more for donating to international relief organizations than donating to domestic organizations.

A number of NGOs facing crackdowns abroad also work on issues that may be seen as "contentious," including human rights, elections, corruption, advocacy, and media freedom—issues that may have a harder time attracting funds due to their non-essential nature when compared to causes like health, education, sanitation, and social services. However, additional education has been shown to increase abstract thinking [@wiepkingmaas, 1978], which may be necessary to donate to causes that do not have quantifiable goals. This is particularly important for organizations with more "contentious" causes as part of their missions [@Bush:2015].

@@

Finally, recent research looks at how a sense of ethnonationalism and national group attachment affects the decision to donate to domestic and international charities. Looking at UK respondents, @hartrobson2019 find that whereas nationalists have a preference for domestic charities, they are not necessarily averse to donating to international causes, whereas internationalists have a strong preference for donating to causes and groups working abroad.

@@

We explore this question using a conjoint experiment conducted with a nationally representative sample in the U.S. Importantly, because there is so little research about international donations during legal crackdown by host governments, our analysis and results are exploratory and descriptive. Accordingly, we do not test specific hypotheses or theories. However, we do talk about causal effects given that we conduct an experiment that varies organizational characteristics in order to elicit opinions from respondents (i.e. specific attributes and features cause different aggregate effects). Instead, we structure our analysis around broader themes from the existing literature and investigate how donors' (1) demographic characteristics, (2) public affairs knowledge and experience, and (3) levels of trust in government, charitable work, and associational life change their donation preferences across different organizational issue areas, relationships with host governments, and funding sources. Our innovative experimental approach provides rich quantitative details of individual and organizational characteristics simultaneously and reveals unique insights into donor motivations that both confirm and expand previous research.

@@

We use a conjoint experiment to simultaneously study the effects of organizational characteristics and individual attributes on individuals' preferences to donate to a nonprofit organization. Conjoint analysis allow us to elicit donor preferences without respondents explicitly stating their exact preferences. This kind of analysis is commonly used in marketing to reveal consumer preferences for hypothetical products, and has become increasingly popular in political science and public policy [@KnudsenJohannesson:2019]. Because we are interested in exploring the complexity behind the decision to donate, we apply conjoint methods to the study of nonprofits and philanthropy.

@@

@tbl:sample-details provides a summary of the individual characteristics of survey respondents. Because we limited the sample to people who regularly make charitable contributions, most respondents have a favorable opinion of nonprofits and charities, with more than half reporting a 5 or greater level of trust in charities (on a 7-point scale). Moreover, nearly 90% indicated that it is important to trust charities. Additionally, the majority of respondents volunteer at least once a year, donate to charity at least once a year, and give between \$1–\$999 annually. Most rarely attend church, but feel that religion is important. Respondents are also civically engaged (three quarters voted in the last election) and follow national news closely (70% follow the news daily). Respondents are fairly balanced ideologically, with the majority reporting a 3, 4, or 5 on a 7-point scale of liberal–conservative political ideology.

@@

In this paper, we look at which kinds of donors are more or less likely to consider an NGO's issue area, funding sources, and relationships with host governments when deciding to donate. Disentangling the effect of every individual level-characteristics measured in the survey on every combination of organizational characteristics presented in the conjoint experiment is unwieldy and near impossible. Given the organizational attributes presented in our experiment, and the possible individual characteristics that respondents could select, there are billions of possible combinations and iterations of organizational and individual attributes—many of which rarely occur in the actual population of potential donors. To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates simultaneously. To simplify the possible combinations of individual characteristics, we generate 64 distinct personas with attributes that vary along different dimensions of demographics, politics and public affairs experience, and social attitudes (see @tbl:persona-attributes). We also generate 24 simulated organizations that vary by issue area, relationship with host government, and source of funding (see @tbl:organization-attributes).

@@

In general, personas respond better to human rights and refugee assistance organizations and environmental advocacy organizations across both levels of income, though wealthier personas show a slight preference towards human rights and refugee organizations. As we explore below, this is likely because income correlates with interest in public affairs and political activity—individuals who are wealthier arguably have more time and resources to follow the news and be continuously engaged in their communities and in other activist causes. There are no sizable differences in donation preferences by income across different forms of NGO funding—both levels of income tend to prefer organizations funded by government grants over small donors.

@@

While education and religiosity are both important drivers of individual philanthropy, neither are associated with any sizable differences in persona donation patterns. In general, both high school and college graduates *and* personas who rarely attend and who often attend religious services follow the same patterns in preferences shown in @fig:income: personas prefer human rights and refugee relief organizations, organizations funded by government grants, and organizations with friendly relaitonships with their host governments (see @fig:education and @fig:religiosity in the appendix.)

@@

@fig:issue-social and @fig:relationship-social show the results of the market simulation across these three dimensions of persona characteristics, varying public affairs knowledge, political ideology, and social trust. Personas respond differently to organizational issue areas. In @fig:issue-social, those who follow the news and travel abroad are more likely to prefer donating to human rights organizations than their less knowledgeable and less traveled counterparts. Personas who do not follow the news or travel show a greater preference for emergency response and refugee relief organization, similar to the income effects seen previously. There are no strong ideological effects across issue areas. Both conservative and liberal personas tend to follow similar patterns of preference across NGO issue.

@@

Personas across both levels of social trust are most likely to donate to donate to organizations with friendly relationships with their host governments. However, those with higher trust are less likely to be deterred by poor host-country relationships, while those with low trust tend to not donate to organizations facing criticism or crackdown abroad. A friendly relationship with a host government is a manifestation of a belief that charities behave properly and exert a positive influence on society. Once an organization faces criticism or crackdown, however, support deteriorates rapidly among personas with low social trust, since such a crackdown could be perceived as a violation of the persona's trust in the charity—that is, the organization must be doing something wrong to deserve criticism. Personas with high levels of trust, on the other hand, seem to stick with legally besieged organizations and show them the benefit of the doubt. The global crackdown on NGOs thus sours NGOs' least trusting donors.

@@

Given the nature of conjoint experiments, the randomly presented organizations may have influenced respondent choices. All four organizations are well known, and it is likely that participant responses reflect prior beliefs and news coverage about Amnesty, Oxfam, Greenpeace, and the Red Cross. For instance, in February 2018, news broke about sexual abuse and misconduct at Oxfam and several other well known organizations, including Amnesty International and Save the Children. This potentially strengthened a narrative that the nonprofit sector suffers from governance failures [@Phillips:2019]. As such, it could be argued that respondents may be less likely to select Oxfam or Amnesty because of these scandals.

However, we remain doubtful that the above news would impact our U.S. respondents' beliefs about these four organizations, as evidence about the effect of these scandals is mixed. Using data from Twitter, @ScurlockDolsakPrakash:2020 find that while Oxfam's reputation suffered for at least six months after the scandal (but has since recovered), other NGOs facing scandals, such as Save the Children, were able to recover more quickly. More broadly, the effect of scandals was mixed—while reputational damage can lead to decline in donations and grants, organizations can become more resilient and durable after a public relations recovery.

Additionally, we remain skeptical that this scandal affected U.S. donors, given Oxfam is based in the UK and lack of awareness about INGOs is well-documented (CITE). Rather, lack of trust towards overseas NGOs, particularly aid and developmental agencies often results due to donors perceiving INGOs as having low personal relevance [@FaulknerTruongRomaniuk:2015]. We failed to find any systematic evidence that the Oxfam scandal influenced perceptions of the nonprofit sector, likely because the story had far less news coverage in the U.S.  Many individual donors have a preference to donate locally and they lack sufficient knowledge to make an informed decision about donating internationally. Overall, the literature suggests that for INGO fundraising, lack of awareness is a much larger issue than specific scandals or other knowledge of specific organizational characteristics that might reduce donor trust and confidence.

The survey conducted among a random sample of U.S. residents with a prior history of charitable giving, also raises questions about generalizability external validity. Public expectations about the role of the public, private, and nonprofit sectors differ by country. In the U.S., for instance, NGOs often receive substantial funding through government grants, while those in the UK and France, such as Amnesty International and Médecins Sans Frontières, receive minimal government funding [@Stroup:2012]. This could imply that donors in UK and France may have very different preferences when choosing to donate to government or privately-funded organizations. As such, more comparative research is needed to see if our results would also replicate cross-nationally.

There is also notable country-based variation in general patterns of philanthropy. Though charitable giving is highest in the U.S, such donations are overwhelmingly directed at local and religious causes. This pattern does not hold for many European countries. In the UK, international aid is one of the largest targets of giving (26%). Other countries prioritize international philanthropy even more, such as Germany (74%), Belgium (61%), and Switzerland (43%) [@Milner:2017]. Thus, if crackdown on INGOs abroad can elicit the effects found in this paper among U.S. respondents, the results provide much reason for optimism for pursuing individual philanthropy towards INGOs in Europe as well. However, more comparative research is required in order to confirm these trends.

@@

More importantly, we find that trust in political institutions and in charitable organizations matters substantially in shaping donor preferences. Personas with high levels of social and charitable trust prefer donating to human rights organizations that have friendly relationships with their host governments, while personas with low trust are drawn more to emergency response and refugee relief organizations with unfriendly relationships with their host governments. Personas with low social trust tend to be scared off by negative host-country relationships and strongly prefer to donate to organizations that are friendly with their host governments, while personas with high social trust are more willing to stick with legally besieged organizations. This could imply that donors that are the most trusting and involved could remain supportive when charities face government criticism and crackdown.

The above results have important implications for NGOs as they navigate a shrinking civic space and consider the use of different frames to tailor their fundraising appeals to individual donors. Our survey finds that at least for some kinds of individuals, NGOs may benefit from publicizing when they are targets of government crackdown, including that information in their fundraising appeals. **TODO THIS IS DIFFERENT NOW** However, the bigger challenge may be addressing concerns of donors with low levels of trust. More broadly, studies from a range of countries show that levels of giving are lowest for refugee and asylum organizations, as well as for overseas and development organizations. Rather, donors giving to international causes tend to support international disaster relief and religious groups [@RobsonHart:2020]. However, the former are precisely the organizations facing crackdown. More research is needed to answer how these organizations can appeal to broader array of potential foreign donors to increase levels of charitable-giving towards them.

@@

# Appendix: Survey experiment

```{r child = "experiment.md"}
```

@@

[^model-details]: See the appendix for technical details about the model and simulation.

>>>>

---
title: "Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy"
short-title: "Who Cares About Crackdowns?"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
date: "September 29, 2020"
published: Working paper.
git-repo: https://example.com
abstract: >-
  Over the past two decades, international non-governmental organizations (INGOs) have faced increasing legal scrutiny and repression abroad. As a result, official aid to INGOs in repressive countries has reduced. NGOs thus need access to a diverse array of funds and may need to increase fundraising from private donors. However, research on the determinants of philanthropy towards international causes—particularly towards organizations working in challenging environments—is still in its infancy. Which demographic and experiential factors influence the decision of individuals to donate to legally besieged NGOs? We investigate this question using a market simulation based on a nationally representative conjoint experiment conducted on individuals across the U.S. with a history of charitable giving. We find that trust in political institutions and charities is enormously influential in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to human rights INGOs with friendly relationships with host governments, and that support holds steady if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do not face criticism or crackdown abroad. Our findings have practical implications for INGOs raising funds from individuals amid closing civic space.
reference-section-title: References
toc: false
mainfont: Lora
sansfont: IBM Plex Sans
---

@@

As a result of these legal crackdowns, foreign aid channeled through NGOs has decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While this has negatively impacted funding flows from public organizations like the United States Agency for International Development (USAID) and the Organisation for Economic Cooperation and Development (OECD), philanthropy from private donors and foundations is not as adversely affected [@McGill:2018]. Given this trend of closing civic space, INGOs need to maintain access to multiple sources of funding, including individual donors.

Individual private donations are a critical source of INGO funding. Total private philanthropy in the United States has grown substantially over time, reaching more than \$400 billion in 2018 (@fig:giving-aggregate). In 2018, private giving to NGOs working in international affairs reached \$22.08 billion, and international affairs-focused organizations are one of only two categories of organizations that have continued experiencing substantial growth in giving [@givingusa:2019].[^intlaffairs] Most of this growth is driven by individuals making small-scale or micro-donations.[^smalldonors] In 2017, 70% of donations to international affairs NGOs came from individuals, rising to 86% in 2018 [@givingusa:2019]. But relying on individual donors means that NGOs must consider individual preferences when fundraising. Which individual donor characteristics and attributes are most salient for INGOs interested in maximizing philanthropy at a time when they are facing increasingly hostile environments abroad? What kinds of donors consider an INGO's host country political environment when making the decision to donate?

@@

There exists a robust scholarship on the determinants of individual-level philanthropy to domestic causes. But the dynamics of *international* giving to NGOs are under-examined in the literature on individual donor attributes and preferences.[^exceptions] Motivating individuals to donate to international causes is hard as the number of recipients are large and far removed from the donor [@CasaleBaumann:2015, 100]. And research on the determinants of philanthropy for international causes—particularly towards organizations working in challenging environments—is still in its infancy due to the absence of expansive and high-quality data [@greenhill2013]. While previous work has looked at how organizational attributes of nonprofits affect donors' willingness to donate internationally [@ChaudhryHeiss:2020], we have yet to uncover demographic traits and experiences that motivate or deter individuals to donate during shrinking civic space.

We explore this question using a simulated market for philanthropic donations, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing to generate dozens of distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that specific individual characteristics help shape donor preferences for donating to INGOs, and that an organization's target issue area and its relationship with its host government interact strongly with these characteristics. Personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with more contentious relationships with host governments, preferring instead to donate to NGOs with friendly host-country relationships and working on less contentious issues like emergency response and refugee relief. Personas with higher trust are more comfortable donating to legally besieged INGOs.

This study contributes both to scholarship and practice in public policy, nonprofits, and philanthropy. Individual giving abroad can help bridge socioeconomic disparities worldwide, especially given the stark inequality in the global distribution of income. Knowledge of demographic traits and its impact on funding can help nonprofits craft more effective and targeted marketing strategies [@Rajan:2009, 414]. Most NGOs working internationally have previously relied on government and foundation funding. But with shrinking civic space, these groups may need to readjust their fundraising strategies towards individuals. Framing is important in this regard—how people process information affects their willingness to donate.[^framing] However, existing research is unclear about which frames are effective when making appeals to individual donors considering international philanthropy. Since our study looks at multiple organizational and host country factors, our results can help NGOs frame their appeals.

@@

Below, we first examine the state of the existing literature on the key drivers of individual giving. We then describe our survey experiment and simulation methods and explore the results. We conclude by discussing the implications of our findings for INGOs facing constrained space for civil society.

@@

Foreign aid to civil society—and NGOs in particular—has steadily increased over time. Many government and foundation donors perceive NGOs as nimbler and less bureaucratic than governments. In countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency, donors specifically seek out NGOs for projects [@Dietrich2013]. However, a restricted NGO environment merits examination of the diverse array of funds available to these organizations.

@@

The advent of funding restrictions on NGOs globally and a subsequent decline in government and foundation funding has made the role and value of individual philanthropy in sustaining NGOs even clearer.[^phildef] NGOs need access to a diverse array of funds and individual donations to international affairs organizations have steadily increased over the past two decades. Moreover, in the United States, donations from individuals have far outstripped foundation and corporate giving. In 2018, individual donations amounted to \$292 billion, compared to \$76 billion from foundations and \$20 billion from corporations [@NPT]. Given that a majority of research is restricted to elite high net-worth donors[^highnetworth] [@USTrust2014], these figures show that it is imperative to understand motivations and attributes of small-scale donors as well.

@@

A vast literature on philanthropy looks at organizational-level factors that motivate individuals to donate [@TBPrakash:2017]. However, donation decisions are not determined solely by a recipient organization's characteristics. Donors have personality traits, preferences, and experiences that make them more or less likely to engage in philanthropy. Substantial research has looked at the determinants of donating to local—rather than international—causes, including four main attributes: social capital, the role of higher education, levels of religiosity, and national group attachment.

@@

Higher education also influences charitable-giving decisions as it fosters prosocial motivations. Education brings people into social networks that entail a higher level of solicitation, which is vital in giving to international causes [@BekkersWiepking:2011]. Education also helps determine the kinds of causes that people donate to globally. @Micklewright:2009 show that higher education matters more for donating to international relief organizations than donating to domestic organizations.

A number of NGOs facing crackdowns abroad also work on issues that may be seen as "contentious," including human rights, elections, corruption, advocacy, and media freedom—issues that may have a harder time attracting funds due to their non-essential nature when compared to causes like health, education, sanitation, and social services [@Bush:2015]. However, additional education has been shown to increase abstract thinking [@wiepkingmaas, 1978], which may be necessary to donate to causes that do not have quantifiable goals, including politically sensitive or contentious causes mentioned above.

@@

Finally, recent research looks at how a sense of ethnonationalism and national group attachment affects the decision to donate to domestic and international charities. Looking at UK respondents, @hartrobson2019 find that whereas nationalists have a preference for domestic charities, they are not necessarily averse to donating to international causes, whereas internationalists have a strong preference for donating to groups working abroad.

@@

We explore this question using a conjoint experiment conducted with a nationally representative sample in the U.S. Importantly, because there is so little research about international donations during legal crackdown by host governments, our analysis and results are exploratory and descriptive. Accordingly, we do not test specific hypotheses or theories. However, we do talk about causal effects given that we conduct an experiment that varies a number of factors in order to elicit opinions from respondents. Instead, we structure our analysis around broader themes from the existing literature and investigate how donors' (1) demographic characteristics, (2) public affairs knowledge and experience, and (3) levels of trust in government, charitable work, and associational life change individual donation preferences across different organizational issue areas, relationships with host governments, and funding sources. Our innovative experimental approach provides rich quantitative details of individual and organizational characteristics simultaneously and reveals unique insights into donor motivations that both confirm and expand previous research.

@@

We use a conjoint experiment to study the simultaneous effects of organizational characteristics and individual attributes on individuals' preferences to donate to a nonprofit organization. Conjoint analysis allows us to elicit donor preferences without respondents explicitly stating their exact preferences. This kind of analysis is commonly used in marketing to reveal consumer preferences for hypothetical products, and has become increasingly popular in political science and public policy [@KnudsenJohannesson:2019]. Because we are interested in exploring the complexity behind the decision to donate, we apply conjoint methods to the study of nonprofits and philanthropy.

@@

@tbl:sample-details provides a summary of the individual characteristics of survey respondents. Because we limited the sample to people who regularly make charitable contributions, most respondents have a favorable opinion of nonprofits and charities, with more than half reporting a 5 or greater level of trust in charities (on a 7-point scale). Moreover, nearly 90% indicated that it is important to trust charities. Additionally, the majority of respondents volunteer and donate to charity at least once a year and give between \$1–\$999 annually. Most rarely attend church, but feel that religion is important. Respondents are also civically engaged (three quarters voted in the last election) and follow national news closely (70% follow the news daily). Respondents are fairly balanced ideologically, with the majority reporting a 3, 4, or 5 on a 7-point scale of liberal–conservative political ideology.

@@

In this paper, we look at which kinds of donors are more or less likely to consider an NGO's issue area, funding sources, and relationships with host governments when deciding to donate. Disentangling the effect of every individual level-characteristic measured in the survey on every combination of organizational characteristics presented in the conjoint experiment is unwieldy and near impossible. Given the organizational attributes presented in our experiment, and the possible individual characteristics that respondents could select, there are billions of possible combinations and iterations of organizational and individual attributes—many of which rarely occur in the actual population of potential donors. To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates simultaneously. To simplify the possible combinations of individual characteristics, we generate 64 distinct personas with attributes that vary along different dimensions of demographics, politics and public affairs experience, and social attitudes (see @tbl:persona-attributes). We also generate 24 simulated organizations that vary by issue area, relationship with host government, and source of funding (see @tbl:organization-attributes).

@@

In general, regardless of income, personas respond better to human rights and refugee assistance organizations and environmental advocacy organizations, followed by emergency response organizations—though wealthier personas show a slight preference towards human rights and refugee organizations. As we explore below, this is likely because income correlates with interest in public affairs and political activity—individuals who are wealthier arguably have more time and resources to follow the news and be continuously engaged in their communities and in other activist causes. There are no sizable differences in donation preferences by income across different forms of NGO funding—both levels of income tend to prefer organizations funded by government grants over small donors.

@@

While education and religiosity are both important drivers of individual philanthropy, neither are associated with any sizable differences in persona donation patterns. In general, both high school and college graduates *and* personas who both rarely and often attend religious services follow the same patterns in preferences shown in @fig:income: personas prefer human rights and refugee relief organizations, organizations funded by government grants, and organizations with friendly relationships with their host governments (see @fig:education and @fig:religiosity in the appendix.)

@@

@fig:issue-social and @fig:relationship-social show the results of the market simulation across these three dimensions of persona characteristics, varying public affairs knowledge, political ideology, and social trust. Personas respond differently to organizational issue areas. In @fig:issue-social, those who follow the news and travel abroad are more likely to prefer donating to human rights organizations than their less knowledgeable and less traveled counterparts. Personas who do not follow the news or travel show a greater preference for emergency response and refugee relief organizations. There are no strong ideological effects across issue areas. Both conservative and liberal personas tend to follow similar patterns of preference across NGO issue.

@@

Personas across both levels of social trust are most likely to donate to organizations with friendly relationships with their host governments. However, those with higher trust are less likely to be deterred by poor host-country relationships, while those with low trust tend to not donate to organizations facing criticism or crackdown abroad. A friendly relationship with a host government may be a manifestation of a belief that charities behave properly and exert a positive influence on society. Once an organization faces criticism or crackdown, however, support deteriorates rapidly among personas with low social trust, since such a crackdown could be perceived as a violation of the persona's trust in the charity—that is, the organization must be doing something wrong to deserve criticism. Personas with high levels of trust, on the other hand, seem to stick with legally besieged organizations and show them the benefit of the doubt. The global crackdown on NGOs thus sours NGOs' least trusting individual donors.

@@

Given the nature of conjoint experiments, the randomly presented organizations may have influenced respondent choices. All four organizations are well-known, and it is likely that participant responses reflect prior beliefs and news coverage about Amnesty, Oxfam, Greenpeace, and the Red Cross. For instance, in February 2018, news broke out about sexual abuse and misconduct at Oxfam and several other prominent organizations, including Amnesty International and Save the Children. This potentially strengthened a narrative that the nonprofit sector suffers from governance failures [@Phillips:2019]. As such, it could be argued that respondents may be less likely to donate to Oxfam or Amnesty because of these scandals.

However, we remain doubtful that the above news would impact our U.S. respondents' beliefs about these four organizations, as evidence about the effect of these scandals is mixed. Using data from Twitter, @ScurlockDolsakPrakash:2020 find that while Oxfam's reputation suffered for at least six months after the scandal (but has since recovered), other NGOs facing scandals, such as Save the Children, were able to recover more quickly. More broadly, the effect of scandals was mixed—while reputational damage can lead to a decline in donations and grants, organizations can become more resilient and durable after a public relations recovery.

Additionally, we remain skeptical that this scandal affected U.S. donors, given Oxfam is based in the UK, and "media narratives are often not as unsupportive as often assumed" [@BanksSchulpenBrockington:2020, 703]. Rather, lack of trust towards overseas NGOs, particularly aid and developmental agencies often results due to donors perceiving INGOs as having low personal relevance [@FaulknerTruongRomaniuk:2015]. We failed to find any systematic evidence that the Oxfam scandal influenced perceptions of the nonprofit sector in the U.S., likely because the story had far less news coverage in the U.S. Many individual donors have a preference to donate locally and they lack sufficient knowledge to make an informed decision about donating internationally. Overall, the literature suggests that for INGO fundraising, lack of awareness is a much larger issue than specific scandals or other knowledge of specific organizational characteristics that might reduce donor trust and confidence.

The survey conducted among a random sample of U.S. residents with a prior history of charitable giving also raises questions about generalizability and external validity. Public expectations about the role of government and private funding of the nonprofit sectors differ by country. In the U.S., for instance, NGOs often receive substantial funding through government grants, while those in the UK and France, such as Amnesty International and Médecins Sans Frontières, receive minimal government funding [@Stroup:2012]. This could imply that donors in UK and France may have very different preferences when choosing to donate to government or privately-funded organizations. More comparative research is needed to see if our results would also replicate cross-nationally.

There is also notable country-based variation in general patterns of philanthropy. Though charitable giving is highest in the U.S, such donations are overwhelmingly directed at local and religious causes. This pattern does not hold for many European countries. In the UK, international aid is one of the largest targets of giving (26%). Other countries prioritize international philanthropy even more, such as Germany (74%), Belgium (61%), and Switzerland (43%) [@Milner:2017]. Thus, if crackdown on INGOs abroad can elicit the effects found in this paper among U.S. respondents, the results provide much reason for optimism for pursuing individual philanthropy towards INGOs in Europe as well. However, more comparative research is required to confirm these trends.

@@

More importantly, we find that trust in political institutions and in charitable organizations matters substantially in shaping donor preferences. Personas with high levels of social and charitable trust prefer donating to human rights organizations that have friendly relationships with their host governments, while personas with low trust are drawn more to emergency response and refugee relief organizations with unfriendly relationships with their host governments. Personas with low social trust tend to be wary of negative host-country relationships and strongly prefer to donate to organizations that are friendly with their host governments, while personas with high social trust are more willing to stick with legally besieged organizations. This could imply that donors that are the most trusting and involved could remain supportive when charities face government criticism and crackdown. The global crackdown on NGOs may also possibly sour NGOs' least trusting individual donors.

The above results have important implications for NGOs as they navigate a shrinking civic space and consider the use of different frames to tailor their fundraising appeals to individual donors. Our survey finds that at least for some kinds of individuals, NGOs may benefit from publicizing when they are targets of government crackdown, including that information in their fundraising appeals. However, the larger challenge may be addressing concerns of donors with low levels of trust who are the least likely to donate organizations facing criticism and crackdown abroad. More broadly, studies from a range of countries show that levels of giving are lowest for refugee and asylum organizations, as well as for overseas and development organizations. Rather, donors giving to international causes tend to support international disaster relief and religious groups [@RobsonHart:2020]. However, the former are precisely the organizations facing crackdown. More research is needed to answer how these organizations can appeal to broader array of potential foreign donors to increase levels of charitable-giving towards them.

@@

@@

[^model-details]: Specifically,

    $$
    \begin{aligned}
    \beta &\sim \operatorname{Multivariate} \mathcal{N}(Z \Gamma, \xi) \\
    y &\sim \operatorname{Multinomial\ logit}(X \beta, \varepsilon)
    \end{aligned}
    $$

    where $y$ = which alternative the respondent chooses to donate to, $X$ = design matrix of attribute levels (i.e. organizations, issue areas, organizational practices, funding sources, and government relations), $\beta$ = latent individual preferences for the attribute levels, $Z$ = matrix of individual-level covariates (i.e. demographics, political knowledge, attitudes towards charity, etc.), $\Gamma$ = matrix of coefficients mapping individual-level covariates onto the latent individual-level preferences, and $\varepsilon$ and $\xi$ = errors.
